---
title: "Chapter 4 : Resampling"
author: "Newton's three sisters"
date: "March 2, 2025"
output:
  beamer_presentation:
    latex_engine: xelatex
    theme: metropolis
  slidy_presentation: default
fonttheme: serif
fontsize: 8pt
institute: Department of Statistics \newline Sungshin Women’s University
header-includes: \input{header_includes.tex}
---

# Cross-Validation

## Cross-Validation

- Assuming that $k$ is an integer that divides $N$, $1/k$ of data are used for the test, and the other $1-k/1$ of the data are used to estimate the model.

<!-- 테이블 -->

   $$ $$   | Group 1  | Group 2  | $\cdots$ | Group $k-1$ | Group $k$
 ----------|----------|----------|----------|-------------|----------
 First     | Test     | Estimate | $\cdots$ | Estimate    | Estimate
 Second    | Estimate | Test     | $\cdots$ | Estimate    | Estimate
  $$ $$    | $\vdots$ | $\vdots$ | $\ddots$ | $\vdots$    | $\vdots$
 $(k-1)$th | Estimate | Estimate | $\cdots$ | Test        | Estimate
 $k$th     | Estimate | Estimate | $\cdots$ | Estimate    | Test
 

## Example with R code
- making linear regression
```{r}
cv.linear= function(X, y, k){
  n = length(y)
  m = n/k # k needs to divide n
  S = 0
  for (j in 1:k){
    test = ((j-1)*m+1):(j*m)
    # specify which out of the n pairs are used for test
    beta = solve(t(X[-test,])%*%X[-test,])%*%t(X[-test,])%*%y[-test]
    # estimate beta using data other than those used for test
    e = y[test] - X[test,]%*%beta
    S = S + drop(t(e)%*%e)
    # evaluate the coefficient beta using data for test
  }
  return(S/n)
}
```

## Apply data set

```{r}
# Data generation
n = 100
p = 5
X = matrix(rnorm(n*p), ncol = p)
X = cbind(1, X)
beta = rnorm(p + 1)
beta[c(2, 3)] = 0
eps = rnorm(n)
y = X%*%beta + eps

# Evaluation via Cross - Validation
cv.linear(X[, c(1, 4, 5, 6)], y, 10)
cv.linear(X[, c(1, 2, 3, 4)], y, 10)
cv.linear(X, y, 10)
```

## Drawing plot
