---
title: "Linear Regression"
author: "Newton's three sisters"
date: "February 9, 2025"
output:
  beamer_presentation:
    latex_engine: xelatex
    theme: metropolis
  slidy_presentation: default
fonttheme: serif
fontsize: 8pt
institute: Department of Statistics \newline Sungshin Women’s University
header-includes: \input{header_includes.tex}
---


# Least Squares Method

## d
```{r warning=FALSE}

```

## c 
```{r}

```

## b
```{r fig.height=5, fig.width=10, message=FALSE}

```

# Multiple Regression

# Distribution $\hat{\beta}$

# Distribution of RSS Values

## Hat matrix
- Hat matrix defined by 


  $\hat{y} = X\hat{\beta} = X(X^TX)^{-1}{X^T}y = Hy$
$$H \triangleq X(X^TX)^{-1}{X^T}$$
- some properties
\begin{align*}
H^2 &= X(X^TX)^{-1}{X^T} \cdot X(X^TX)^{-1}{X^T} = X(X^TX)^{-1}{X^T} = H \\
(I -H)^2 &= I - 2H + H^2 = I - H \\
HX &= X(X^TX)^{-1}{X^T} \cdot X = X
\end{align*}

## Residual sum of square
- RSS defined
$$
\text{RSS} \triangleq ||y-\hat{y}||^2
$$
- Using hat matrix
\begin{align*}
y - \hat{y} &= y - Hy = (I-H)y = (I-H)(X\beta+\varepsilon) \\
            &= (X -HX)\beta + (I - H)\varepsilon = (I - H)\varepsilon
\end{align*}
$$
RSS \triangleq ||y - \hat{y}||^2 = \{(I - H)\varepsilon\}^T (I - H)\varepsilon = \varepsilon^T(I-H)^2\varepsilon = \varepsilon^T(I-H)\varepsilon
$$

## Eigenvalues of $H$ and Null space of ($I-H$)
- Proof by contrapositive
\begin{align*}
Hx = x \Rightarrow (I - H)x = 0 \\
(I - H)x = 0 \Rightarrow Hx = x
\end{align*}
- Dimensions of the eigenspaces of $H$ is $p + 1$ \
\vt
\textbf{Proof} using rank($X) = p+1$ 
\begin{align*}
\text{rank}(H) &\le \text{min}\{\text{rank}(X(X^TX)^{-1}), \text{rank}(X)\} \le \text{rank}(X) = p + 1 \\
\text{rank}(H) &\ge \text{rank}(HX) = \text{rank}(X) = p + 1
\end{align*}

- Dimensions of the null space of $I-H$ is $N - (p+1)$

$$
P (I - H) P^T = \operatorname{diag}(\underbrace{1, \dots, 1}_{N-p-1}, \underbrace{0, \dots, 0}_{p+1})
$$

## 제목 뭐라고 하지..
-  We define $v = P\varepsilon \in \mathbb{R}^N$, then from $\varepsilon = P^Tv$
\begin{align*}
\text{RSS} 
&= \varepsilon^T(I - H)\varepsilon = (P^Tv)^T(I-H)P^Tv = v^TP(I-H)P^Tv \\
&= [v_1, \cdots, v_{N-p-1}, v_{N-p}, \cdots, v_n] 
\begin{bmatrix}
1      & 0      &\cdots  &  \cdots & \cdots & 0 \\
0      & \ddots & 0      &  \cdots & \cdots & \vdots \\
\vdots & 0      & 1      &  \cdots & \cdots & 0 \\
0      & 0      & 0      &  \cdots & \cdots & \cdots \\
\vdots & \vdots & \vdots &  \vdots & \ddots & \vdots \\
0      & \cdots & 0      &  \cdots & \cdots & 0
\end{bmatrix}
\begin{bmatrix}
v_1 \\
\vdots \\
v_{N-p-1} \\
v_{N-p} \\
\vdots \\
v_N
\end{bmatrix} 
&= \sum^{N-p-1}_{i = 1}v_i^2
\end{align*}

## dd
- Let $w \in \mathbb{R}^{N-p-1}$
\vt
- Average 

  $E[v] = E[P\varepsilon] = 0$

  $E[w] = 0$
\vt

- Covariance 

  $E[vv^t] = E[P\varepsilon(P\varepsilon)^T] = PE[\varepsilon\varepsilon^t]P=P\sigma^2IP^T = \sigma^2I$
  
  $E[ww^T] = \sigma^2I$
\vt

- We have RSS
$$
\frac{RSS}{\sigma^2} \sim \chi^2_{N-p-1}
$$

# Hypothesis Testing for $\hat{\beta}_j \ne 0$

# Coefficient of Determination and the Detection of Collinearity

# Confidence and Prediction Intervals

## a
```{r}

```
<!-- ##  -->
<!-- ```{r message=FALSE} -->
<!-- shinyApp(ui, server) -->
<!-- ``` -->


## Q & A

\begin{center}
  {\bf {\Huge Q \& A}}
\end{center}

## 

\begin{center}
  {\bf {\Huge Thank you :)}}
\end{center}

