---
title: "Linear Regression"
author: "Newton's three sisters"
date: "February 9, 2025"
output:
  beamer_presentation:
    latex_engine: xelatex
    theme: metropolis
  slidy_presentation: default
fonttheme: serif
fontsize: 8pt
institute: Department of Statistics \newline Sungshin Women’s University
header-includes: \input{header_includes.tex}
---


# Least Squares Method

## d
```{r warning=FALSE}

```

## c 
```{r}

```

## b
```{r fig.height=5, fig.width=10, message=FALSE}

```

# Multiple Regression

# Distribution $\hat{\beta}$

# Distribution of RSS Values

## Hat matrix
- Hat matrix defined by $\hat{y} = Hy$


  $\hat{y} = X\hat{\beta} = X(X^TX)^{-1}{X^T}y = Hy$
$$H \triangleq X(X^TX)^{-1}{X^T}$$
- some properties
\begin{align*}
H^2 &= X(X^TX)^{-1}{X^T} \cdot X(X^TX)^{-1}{X^T} = X(X^TX)^{-1}{X^T} = H \\
(I -H)^2 &= I - 2H + H^2 = I - H \\
HX &= X(X^TX)^{-1}{X^T} \cdot X = X
\end{align*}

## Residual sum of square
- RSS defined
$$
\text{RSS} \triangleq ||y-\hat{y}||^2
$$
- using hat matrix
\begin{align*}
y - \hat{y} &= y - Hy = (I-H)y = (I-H)(X\beta+\varepsilon) \\
            &= (X -HX)\beta + (I - H)\varepsilon = (I - H)\varepsilon
\end{align*}
$$
RSS \triangleq ||y - \hat{y}||^2 = \{(I - H)\varepsilon\}^T (I - H)\varepsilon = \varepsilon^T(I-H)^2\varepsilon = \varepsilon^T(I-H)\varepsilon
$$

## Eigenvalues $H$ and $I-H$
- They are only zeros and ones
\vt
- Dimensions of the eigenspaces of $H$ and $I-H$ are both $p + 1$ \
\vt
\textbf{Proof} using rank($X) = p+1$ 
\begin{align*}
\text{rank}(H) &\le \text{min}\{\text{rank}(X(X^TX)^{-1}), \text{rank}(X)\} \le \text{rank}(X) = p + 1 \\
\text{rank}(H) &\ge \text{rank}(HX) = \text{rank}(X) = p + 1
\end{align*}

# Hypothesis Testing for $\hat{\beta}_j \ne 0$

# Coefficient of Determination and the Detection of Collinearity

# Confidence and Prediction Intervals

## a
```{r}

```
<!-- ##  -->
<!-- ```{r message=FALSE} -->
<!-- shinyApp(ui, server) -->
<!-- ``` -->


## Q & A

\begin{center}
  {\bf {\Huge Q \& A}}
\end{center}

## 

\begin{center}
  {\bf {\Huge Thank you :)}}
\end{center}

