---
title: "Classification"
author: "Jang Miyoung"
date: "February 9, 2025"
output:
  beamer_presentation:
    latex_engine: xelatex
    theme: metropolis
  slidy_presentation: default
fonttheme: serif
fontsize: 8pt
institute: Department of Statistics \newline Sungshin Women’s University
header-includes: \input{header_includes.tex}
---

# Logistic Regression

## ad

- The probabilities of $y = 1$ and $y = -1$ are expressed by $\frac{e^{\beta_0+x\beta}}{1 + e^{\beta_0+x\beta}}$ and $\frac{1}{1 + e^{\beta_0+x\beta}}$

$$
\frac{1}{1 + e^{-y(\beta_0+x\beta)}}
$$

```{r fig.height=5, fig.width=10, echo=FALSE}
f = function(x){
  exp(beta.0 + beta*x)/(1+exp(beta.0 + beta*x))
}
beta.0 = 0; beta.seq = c(0, 0.2, 0.5, 1, 2, 10); m = length(beta.seq); beta = beta.seq[1]
plot(f, xlim = c(-10,10), ylim = c(0, 1), xlab = "x", ylab = "P(Y=1|x)", col = 1,
     main = "Logistic Curve")
for (i in 2:m){
  beta = beta.seq[i]; par(new = TRUE);
  plot(f, xlim = c(-10, 10), ylim = c(0, 1), xlab = " ", ylab = " ", axes = FALSE, col = i)
}
legend("topleft", legend = beta.seq, col = 1:length(beta.seq), lwd = 2, cex = .8)
```
 
## increasing monotonically and convex and concave

\begin{align*}
f'(x) &= \beta\frac{e^{-(\beta_0+x\beta)}}{(1+e^{-(\beta_0+x\beta)})^2} \ge 0 \\
f''(x) = &-\beta^2\frac{e^{-(\beta_0+x\beta)}[1-e^{-(\beta_0+x\beta)}]}{(1+e^{-(\beta_0+x\beta)})^3}
\end{align*}

- We see that $f(x)$ is increasing monotonically and is convex and concave when $x<-\beta_0/\beta$ and $x>-\beta_0/\beta$, they chage at $x=0$, when $\beta =0$

# Netwon Raphson Method

## Netwon Raphson 

- Tangent line is $y - f(x_i) = f'(x_i)(x - x_i)$ the intersection with y = 0

$$
x_{i+1} \triangleq x_i - \frac{f(x_i)}{f'(x_i)}
$$

- Example $f(x) = x^2-1$ and $x_0 = 4$ 

```{r fig.height=5, fig.width=10, echo=FALSE}
f = function(x){
  x^2-1
}
f. = function(x){
  2*x
}

curve(f(x), -1, 5)
abline(h=0, col = "blue")
x = 4
for (i in 1:10){
  X = x
  Y = f(x)
  x = x - f(x)/f.(x)
  y = f(x)
  segments(X, Y, x, 0)
  segments(X, Y, X, 0, lty = 3)
  points(x, 0, col = "red", pch = 16)
}
```

## Newton Raphson Method for two variables
